### 2024-09-13 main.tex
Updates from Overleaf

% CVPR 2025 Paper Template; see https://github.com/cvpr-org/author-kit
\documentclass[10pt,twocolumn,letterpaper]{article}

%%%%%%%%% PAPER TYPE  - PLEASE UPDATE FOR FINAL VERSION
\usepackage[review]{cvpr} % For the REVIEW version
% \usepackage[pagenumbers]{cvpr} % For the arXiv version
% \usepackage{cvpr} % For the CAMERA-READY version

%%%%%%%%% Import additional packages in the preamble file
\input{preamble}

%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\paperID{*****} % *** Enter the Paper ID here
\def\confName{CVPR}
\def\confYear{2025}

%%%%%%%%% TITLE - PLEASE UPDATE
\title{\LaTeX\ Author Guidelines for \confName~\confYear~Proceedings}

%%%%%%%%% AUTHORS - PLEASE UPDATE
\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

%%%%%%%%% SELECT WHAT TO TYPESET - SELECT ONE
\includeonly{_paper}
% \includeonly{_supplementary}
% \includeonly{_rebuttal}

\begin{document}
\include{_paper}
\include{_supplementary}
\include{_rebuttal}
\end{document}