### 2024-11-10 _supplementary.tex

\clearpage
\maketitleagain{Supplementary Material}

The supplementary can be compiled together with the main paper, therefore:
% 
\begin{itemize}
\item The supplementary can back-reference sections of the main paper, for example, we can refer to \cref{sec:intro};
\item The main paper can forward reference sub-sections within the supplementary explicitly (e.g. referring to a particular experiment); 
\item When submitted to arXiv, the supplementary will already included at the end of the paper.
\item To submit the supplementary, use the \verb|\includeonly{| command to avoid typesetting the paper itself.
\item You can also add citations that you may have missed~\cite{Alpher2005}.
\end{itemize}

{
    \small
    \bibliographystyle{ieeenat_fullname}
    \bibliography{main.bib}
}